The Near and Mid-infrared photometric properties of known redshift $z\geq5$ Quasars

### 2020-01-09 tex/MNRAS_submission/rv1/VHzQ_MNRAS_rv1.tex
VHzQ_MNRAS_rv1.tex updated

\maketitle
\begin{abstract}
We assemble a catalogue of 463 spectroscopically confirmed very high
($z\geq5.00$) redshift quasars and report their near- ($ZYJHK_{s}/K$)
and mid- (WISE W1234) infrared properties. 97\% of the VH$z$Q sample
is detected in one or more NIR ($ZYJHK/K_{s}$) band, with lack of
coverage rather than lack of depth being the reason for the
non-detections. 362 (78\%) of the very high redshift quasars are
detected at 3.4$\mu$m in the W1 band from the unWISE catalog and all
of the $z\geq7$ quasars are detected in both unWISE W1 and W2. Using
archival WFCAM/UKIRT and VIRCAM/VISTA data we check for photometric
variability that might be expected from super-Eddington accretion. We
find 32 of the quasars have sufficient NIR measurements and
signal-to-noise to look for variability. Weak variability was detected
in multiple bands of SDSS J0959+0227, and very marginally in the
$Y$-band of MMT J0215-0529. Only one quasar, SDSS J0349+0034, shows
significantdifferences between WFCAM and VISTA magnitudes in one band.
With supermassive black hole accretion likely to be redshift invariant
up to very high-redshift, further monitoring of these sources is
warranted.  All the data, analysis codes and plots used and generated
here can be found at:
\href{https://github.com/d80b2t/VHzQ}{\tt github.com/d80b2t/VHzQ}.
\end{abstract}

@@

This paper is organized as follows. In Section 2, we present the
assembled list of the 463 $z\geq5.00$ VH$z$Qs that we have
compiled. We then give a high-level overview of the photometric
surveys and datasets we use and present the photometry of the
VH$z$Qs. In Section 3 we investigate the variability properties of the
VH$z$Qs, looking for evidence of super-critical accretion. In Section
4 we calculate how many $z>5$ quasars we should expect to find in
current datasets. We conclude in Section 5 and present all the
necessary details to obtain our dataset in the Appendices.

@@

We have compiled a list of 463 quasars with redshifts $z\geq5.00$. We
use all the $z\geq5.00$ quasars that have been discovered,
spectroscopically confirmed and published as of 2018 December 31 (MJD
58483). We then obtain optical, near-infrared and mid-infrared
photometry for the spectral dataset. The near-infrared data comes from
two sources: first, the WFCAM \citep[][]{Casali2007} on the UKIRT,
primarily, but not exclusively, as part of the UKIRT Infrared Deep Sky
Survey \citep[UKIDSS; ][]{Lawrence2007}.  And second, data from the
VIRCAM on the VISTA \citep[][]{Emerson2006, Dalton2006}. The
mid-infrared, $\lambda=3-30\mu$m wavelength data is from the 
Wide-Field Infrared Survey Explorer \citep[WISE;][]{Wright2010,
Cutri2013} mission. For reference, Figure~\ref{fig:filters} displays
the wavelength and normalised transmission of the filters in question.

\subsection{Spectroscopy} 
We compile the list of all known, spectroscopically confirmed
quasars from the literature.  This list was complied from a range of
surveys and papers.  Specifically, we use data from:
\citet{Banados2014, Banados2016, Banados2018a, Banados2018b}, 
\citet{Becker2015}, 
\citet{Calura2014}, 
\citet{Carilli2007, Carilli2010}, 
\citet{Carnall2015}, 
\citet{Cool2006}, 
\citet{Douglas2007}, 
\citet{DeRosa2011}, 
\citet{Fan2000, Fan2001c, Fan2003, Fan2004, Fan2006, Fan2018}, 
\citet{Goto2006}, 
\citet{Ikeda2017}, 
\citet{Jiang2008, Jiang2009, Jiang2015, Jiang2016},   
\citet{Kashikawa2015}, 
\citet{Khorunzhev2017}, 
\citet{Koptelova2017}, 
\citet{Kozlowski2019}, 
\citet{Kim2015, Kim2018},  
\citet{Kurk2007, Kurk2009}, 
\citet{Leipski2014}, 
\citet{Mahabal2005}, 
\citet{Matsuoka2016,  Matsuoka2018a, Matsuoka2018b},   
\citet{Mazzucchelli2017}, 
\citet{Morganson2012}, 
\citet{Mortlock2009, Mortlock2011},
\citet{McGreer2006, McGreer2013},  
\citet{Reed2015, Reed2017}, 
\citet{Saxena2018}, 
\citet{Stern2007},  
\citet{Tang2017}, 
\citet{vanBreugel1999}, 
\citet{Venemans2007, Venemans2012, Venemans2013, Venemans2015a, Venemans2015b, Venemans2016},
\citet{WangF2016, WangF2017, WangF2018a, WangF2018b},
\citet{Willott2007, Willott2009, Willott2010a, Willott2013b, Willott2015}, 
\citet{Wu2015} 
\citet{YangJ2018a, YangJ2018b}  
and 
\citet{Zeimann2011}.

@@

\begin{table}
\centering
\begin{tabular}{l r  r}
\hline \hline
$z \geq$  & Age / Myr & No. of objects \\
\hline 
5.00         & 1180          &  463   \\
5.70         & 1000          &  267   \\
6.00         &   937          &  170   \\
6.19         &   900          &  86   \\
6.50         &    845         &  40   \\
6.78         &    800         &  14   \\
7.00         &    767         &   4   \\
7.50         &    700         &   1   \\
\hline \hline
\end{tabular}
\caption{The number of objects at or above a given redshift. 
The age of the Universe in Megayears is also given. }
      \label{tab:ages}
\end{table}

@@

  We query the WSA and VSA performing matched-aperture photometry at
  the positions of our 463 VH$z$Qs. This database is world-readable and
  we give the full recipe and relevant SQL queries for accessing both
  databases in Appendix~\ref{sec:SQL} as well as online.

@@

\section{Results}
Having collated the sample of 463 VH$z$Qs, and obtained their near-
and mid-infrared photometry we report here the various photometric
properties of the quasars.
%%
Table~\ref{tab:output_table} represents the culmination of this
effort, {\it is the main
\href{https://github.com/d80b2t/VHzQ/blob/master/data/VHzQs_ZYJHK_WISE.dat}{data
product} of this paper} and we now exam the assembly of its contents
in more detail.

@@

  \subsubsection{Comparing WFCAM and VISTA}
  There are 130 quasars observed in the overlapping area between WFCAM
  and VISTA. We used the {\tt VegaToAB} values to put these objects on
  the same AB system, and for each object compared the two
  measurements. First, we calculated the weighted average (calibrated
  flux) in each filter of both and calculated the ratio and difference
  between each measurement and the average.  Then for each filter we
  calculated the weighted average of the differences (in mag) for each
  instrument to see if there were significant offsets. The results are
  given in Table~\ref{tab:WFCAM_vs_VISTA}.  The only filter with a
  significant offset is the $Y$-band. All of the VISTA averages are
  negative and all of the WFCAM ones are positive. The $Ks$ versus $K$
  band comparison may be affected the different shapes of the filters,
  with $K$ being significantly wider than $K_s$, able to detect light at
  longer wavelengths.
  \begin{table}
    \centering
    \begin{tabular}{l r r}
      \hline  \hline
      abs(VIRCAM & \multirow{2}{*}{millimags} &  no. of  \\
      -  WFCAM)      &                                        &  objects \\
      \hline
      $Z$                 &  23.2 	& 3 \\
      $Y$                 &  57.3 	& 53 \\
      $J$                  &    2.1 	& 106 \\
      $H$                 &  45.8     &  96 \\
      $K_{\rm s}$/$K$ &  25.2     & 110 \\
      \hline  \hline
    \end{tabular}
    \caption{Comparing the magnitudes in different WFCAM/UKIRT and 
      VIRCAM/VISTA near-infrared bands.}
    \label{tab:WFCAM_vs_VISTA}
  \end{table}

@@

\citet{Secrest2015} present an all-sky sample of $\approx$1.35 million
AGN meeting a two-color, (W1-W2) vs. (W2-W3), photometric selection as
applied to sources from AllWISE. This colour-selection is motivated by
\citet{Mateos2012}. We investigate if this catalogue, which is tuned
for, and noted to be highly complete at, lower, $z\lesssim2$
redshifts, contain the VH$z$Qs. We find only 5 VH$z$Qs from our sample
match the Secrest catalogue, but given a W3 detection is required,
this is not surprising. Further, in addition to the relative lack of
sensitivity in W3, the mid-IR colors are moving out of the
\citet{Mateos2012} color criterion at high redshift. This was shown
using an AGN SED where AGNs above a redshift of about $\sim$2 fall out
of the colour criterion, see e.g. Figure 5 of \citet[][]{Mateos2012}
and Figure 9 of \citet{Secrest2015}.
%%
\citet{Assef2018} released two large catalogues of AGN
candidates identified across 30,000 deg$^2$ of extragalactic sky from
the WISE AllWISE Data Release. The ``R90'' catalogue, contains 4.5M
AGN candidates at 90\% reliability ($\approx$150 AGN candidates
per deg$^2$) while the ``C75'' catalogue consists of 20.9M AGN
candidates at 75\% completeness ($\approx$700 AGN candidates per
deg$^2$). Crossmatching our catalogue of 463 VH$z$Qs with these
catalogues, produces 42 matches with the R90 sample and 98 matches
with the C75 sample. Both catalogues unsurprisingly match to the
ultraluminous quasar SDSS J0100+2802 \citep{Wu2015} while the C75, but
not the R90 catalogue matches to the first redshift $z>7$ quasar ULAS
J1120+0641 \citep{Mortlock2011}. Neither the R90 or C75 catalogue matches to the
current highest-redshift object J1342+0928 \citep{Banados2018}.

@@

We thank:
\begin{list}{$\circ$}{}  
  \item Mike Read at the ROE WFAU for help with the WFCAM Science Archive (WSA) and the VISTA Science Archive (VSA). 
  \item Aaron Meisner and Eddie Schlafly for facilitating early access to the unWISE Catalog. 
  \item Tim Brooke at the \href{https://irsasupport.ipac.caltech.edu/index.php}{IRSA Help Desk}; 
  \item Bernie Shiao at STScI for help with the Pan-STARRS1 DR1 CasJobs interface. 
  \item Michael Cushing for supplying the Late Type stellar spectra and Beth Biller for useful discussion. 
  \item Nathan Secrest for useful discussions on the mid-IR AGN catalogues. 
\end{list}

@@

At the \href{http://horus.roe.ac.uk/vsa/login.html}{VSA Login}, enter 
with these credentials::
\begin{itemize}
    \item Username: {\tt VSERV1000} 
    \item password: {\tt highzqso} 
    \item community: {\tt proprietary}
\end{itemize}
Then the \href{http://horus.roe.ac.uk:8080/vdfs/VSQL_form.jsp}{Freeform SQL
Query} page is where the database release to use is {\tt
VSERV1000v20190508}. The queries are the same as those used in the
WSA, see Appendix~\ref{sec:SQL}, but the names of the tables starting
with {\tt wserv1000} should be replaced with {\tt vserv1000}.

>>>>

\maketitle
\begin{abstract}
We assemble a catalogue of 488 spectroscopically confirmed very high
($z\geq5.00$) redshift quasars and report their near- ($ZYJHK_{s}/K$)
and mid- (WISE W1234) infrared properties. 97\% of the VH$z$Q sample
is detected in one or more NIR ($ZYJHK/K_{s}$) band, with lack of
coverage rather than lack of depth being the reason for the
non-detections. 362 (78\%) of the very high redshift quasars are
detected at 3.4$\mu$m in the W1 band from the unWISE catalog and all
of the $z\geq7$ quasars are detected in both unWISE W1 and W2. Using
archival WFCAM/UKIRT and VIRCAM/VISTA data we check for photometric
variability that might be expected from super-Eddington accretion. We
find 32 of the quasars have sufficient NIR measurements and
signal-to-noise to look for variability. Weak variability was detected
in multiple bands of SDSS J0959+0227, and very marginally in the
$Y$-band of MMT J0215-0529. Only one quasar, SDSS J0349+0034, shows
significantdifferences between WFCAM and VISTA magnitudes in one band.
With supermassive black hole accretion likely to be redshift invariant
up to very high-redshift, further monitoring of these sources is
warranted.  All the data, analysis codes and plots used and generated
here can be found at:
\href{https://github.com/d80b2t/VHzQ}{\tt github.com/d80b2t/VHzQ}.
\end{abstract}

@@

This paper is organized as follows. In Section 2, we present the
assembled list of the 488 $z\geq5.00$ VH$z$Qs that we have
compiled. We then give a high-level overview of the photometric
surveys and datasets we use and present the photometry of the
VH$z$Qs. In Section 3 we investigate the variability properties of the
VH$z$Qs, looking for evidence of super-critical accretion. In Section
4 we calculate how many $z>5$ quasars we should expect to find in
current datasets. We conclude in Section 5 and present all the
necessary details to obtain our dataset in the Appendices.

@@

We have compiled a list of 488 quasars with redshifts $z\geq5.00$. We
use all the $z\geq5.00$ quasars that have been discovered,
spectroscopically confirmed and published as of 2018 December 31 (MJD
58483). We then obtain optical, near-infrared and mid-infrared
photometry for the spectral dataset. The near-infrared data comes from
two sources: first, the WFCAM \citep[][]{Casali2007} on the UKIRT,
primarily, but not exclusively, as part of the UKIRT Infrared Deep Sky
Survey \citep[UKIDSS; ][]{Lawrence2007}.  And second, data from the
VIRCAM on the VISTA \citep[][]{Emerson2006, Dalton2006}. The
mid-infrared, $\lambda=3-30\mu$m wavelength data is from the 
Wide-Field Infrared Survey Explorer \citep[WISE;][]{Wright2010,
Cutri2013} mission. For reference, Figure~\ref{fig:filters} displays
the wavelength and normalised transmission of the filters in question.

\subsection{Spectroscopy} 
We compile the list of all known, spectroscopically confirmed
quasars from the literature.  This list was complied from a range of
surveys and papers.  Specifically, we use data from:
\citet{Banados2014, Banados2016, Banados2018a, Banados2018b}, 
\citet{Becker2015}, 
\citet{Calura2014}, 
\citet{Carilli2007, Carilli2010}, 
\citet{Carnall2015}, 
\citet{Cool2006}, 
\citet{Douglas2007}, 
\citet{DeRosa2011}, 
\citet{Fan2000, Fan2001c, Fan2003, Fan2004, Fan2006, Fan2018}, 
\citet{Goto2006}, 
\citet{Ikeda2017}, 
\citet{Jiang2008, Jiang2009, Jiang2015, Jiang2016},   
\citet{Kashikawa2015}, 
\citet{Khorunzhev2017}, 
\citet{Koptelova2017}, 
\citet{Kozlowski2019}, 
\citet{Kim2015, Kim2018},  
\citet{Kurk2007, Kurk2009}, 
\citet{Leipski2014}, 
\citet{Mahabal2005}, 
\citet{Matsuoka2016,  Matsuoka2018a, Matsuoka2018b},   
\citet{Mazzucchelli2017}, 
\citet{Morganson2012}, 
\citet{Mortlock2009, Mortlock2011},
\citet{McGreer2006, McGreer2013},  
\citet{Reed2015, Reed2017}, 
\citet{Saxena2018},
\citet{Schneider2007}, 
\citet{Stern2007},  
\citet{Tang2017}, 
\citet{vanBreugel1999}, 
\citet{Venemans2007, Venemans2012, Venemans2013, Venemans2015a, Venemans2015b, Venemans2016},
\citet{WangF2016, WangF2017, WangF2018a, WangF2018b},
\citet{Willott2007, Willott2009, Willott2010a, Willott2013b, Willott2015}, 
\citet{Wu2015} 
\citet{YangJ2018a, YangJ2018b}  
and 
\citet{Zeimann2011}.

We updated our list of objects between the submitted paper and accepted due
to feedback from other scientists. We discuss these changes in detail in
Appendix~\ref{app:qso_sample}.

@@

\begin{table}
\centering
\begin{tabular}{l r  r}
\hline \hline
$z \geq$  & Age / Myr & No. of objects \\
\hline 
5.00         & 1180          &  488   \\
5.70         & 1000          &  273  \\
6.00         &   937          &  174   \\
6.19         &   900          &  87   \\
6.50         &    845         &  40   \\
6.78         &    800         &  14   \\
7.00         &    767         &   4   \\
7.50         &    700         &   1   \\
\hline \hline
\end{tabular}
\caption{The number of objects at or above a given redshift. 
The age of the Universe in Megayears is also given. }
      \label{tab:ages}
\end{table}

@@

  We query the WSA and VSA performing matched-aperture photometry at
  the positions of our 488 VH$z$Qs. This database is world-readable and
  we give the full recipe and relevant SQL queries for accessing both
  databases in Appendix~\ref{sec:SQL_WSA} as well as online.

@@

\section{Results}
Having collated the sample of 488 VH$z$Qs, and obtained their near-
and mid-infrared photometry we report here the various photometric
properties of the quasars.
%%
Table~\ref{tab:output_table} represents the culmination of this
effort, {\it is the main
\href{https://github.com/d80b2t/VHzQ/blob/master/data/VHzQs_ZYJHK_WISE.dat}{data
product} of this paper} and we now exam the assembly of its contents
in more detail.

@@

  \subsubsection{Comparing WFCAM and VISTA}
  There are 131 quasars observed in the overlapping area between WFCAM
  and VISTA. We used the {\tt VegaToAB} values to put these objects on
  the same AB system, and for each object compared the two
  measurements. First, we calculated the weighted average (calibrated
  flux) in each filter of both and calculated the ratio and difference
  between each measurement and the average.  Then for each filter we
  calculated the weighted average of the differences (in mag) for each
  instrument to see if there were significant offsets. The results are
  given in Table~\ref{tab:WFCAM_vs_VISTA}.  The only filter with a
  significant offset is the $Y$-band. All of the VISTA averages are
  negative and all of the WFCAM ones are positive. The $Ks$ versus $K$
  band comparison may be affected the different shapes of the filters,
  with $K$ being significantly wider than $K_s$, able to detect light at
  longer wavelengths.
  \begin{table}
    \centering
    \begin{tabular}{l r r}
      \hline  \hline
      abs(VIRCAM & \multirow{2}{*}{millimags} &  no. of  \\
      -  WFCAM)      &                                        &  objects \\
      \hline
      $Z$                 &  7.2 	& 3 \\
      $Y$                 &  59.1 	& 62 \\
      $J$                  &    6.3 	& 111 \\
      $H$                 &  41.7     &  103 \\
      $K_{\rm s}$/$K$ &  15.1     & 114 \\
      \hline  \hline
    \end{tabular}
    \caption{Comparing the magnitudes in different WFCAM/UKIRT and 
      VIRCAM/VISTA near-infrared bands.}
    \label{tab:WFCAM_vs_VISTA}
  \end{table}

@@

\citet{Secrest2015} present an all-sky sample of $\approx$1.35 million
AGN meeting a two-color, (W1-W2) vs. (W2-W3), photometric selection as
applied to sources from AllWISE. This colour-selection is motivated by
\citet{Mateos2012}. We investigate if this catalogue, which is tuned
for, and noted to be highly complete at, lower, $z\lesssim2$
redshifts, contain the VH$z$Qs. We find only 5 VH$z$Qs from our sample
match the Secrest catalogue, but given a W3 detection is required,
this is not surprising. Further, in addition to the relative lack of
sensitivity in W3, the mid-IR colors are moving out of the
\citet{Mateos2012} color criterion at high redshift. This was shown
using an AGN SED where AGNs above a redshift of about $\sim$2 fall out
of the colour criterion, see e.g. Figure 5 of \citet[][]{Mateos2012}
and Figure 9 of \citet{Secrest2015}.
%%
\citet{Assef2018} released two large catalogues of AGN
candidates identified across 30,000 deg$^2$ of extragalactic sky from
the WISE AllWISE Data Release. The ``R90'' catalogue, contains 4.5M
AGN candidates at 90\% reliability ($\approx$150 AGN candidates
per deg$^2$) while the ``C75'' catalogue consists of 20.9M AGN
candidates at 75\% completeness ($\approx$700 AGN candidates per
deg$^2$). Crossmatching our catalogue of 488 VH$z$Qs with these
catalogues, produces 42 matches with the R90 sample and 98 matches
with the C75 sample. Both catalogues unsurprisingly match to the
ultraluminous quasar SDSS J0100+2802 \citep{Wu2015} while the C75, but
not the R90 catalogue matches to the first redshift $z>7$ quasar ULAS
J1120+0641 \citep{Mortlock2011}. Neither the R90 or C75 catalogue matches to the
current highest-redshift object J1342+0928 \citep{Banados2018a}.

@@

We thank:
\begin{list}{$\circ$}{}  
  \item Mike Read at the ROE WFAU for help with the WFCAM Science Archive (WSA) and the VISTA Science Archive (VSA). 
  \item Aaron Meisner and Eddie Schlafly for facilitating early access to the unWISE Catalog.
  \item Krisztina Perger for key discussions about missing objects after the pre-print appeared. 
  \item Tim Brooke at the \href{https://irsasupport.ipac.caltech.edu/index.php}{IRSA Help Desk}; 
  \item Bernie Shiao at STScI for help with the Pan-STARRS1 DR1 CasJobs interface. 
  \item Michael Cushing for supplying the Late Type stellar spectra and Beth Biller for useful discussion. 
  \item Nathan Secrest for useful discussions on the mid-IR AGN catalogues. 
\end{list}

@@

At the \href{http://horus.roe.ac.uk/vsa/login.html}{VSA Login}, enter 
with these credentials::
\begin{itemize}
    \item Username: {\tt VSERV1000} 
    \item password: {\tt highzqso} 
    \item community: {\tt proprietary}
\end{itemize}
Then the \href{http://horus.roe.ac.uk:8080/vdfs/VSQL_form.jsp}{Freeform SQL
Query} page is where the database release to use is {\tt
VSERV1000v20190508}. The queries are the same as those used in the
WSA, see Appendix~\ref{sec:SQL_WSA}, but the names of the tables starting
with {\tt wserv1000} should be replaced with {\tt vserv1000}.

\section{Updated VH$z$Q Sample}
\label{app:qso_sample}
Between the journal submitted version of this paper (which was
also the arXiv v1) we updated the sample of VH$z$Qs from a total of
463 objects to 488 objects.  16 of these objects were from
\cite{YangJ2018a} and 9 objects were added after discussions with
Krisztina Perger, with the relevant catalog found at 
\href{http://astro.elte.hu/$\sim$perger/catalog.html}{astro.elte.hu/$\sim$perger/catalog.html}.
We do not include the object DES0141-54 \citep{Belladitta2018}, since we were unable to 
obtain arcsecond level coordinates to perform our matched-aperture photometry.
We also update the positions of five objects, see Table~\ref{tab:disagree_objs}.

One object in the Perger catalogue, VA-13492 we could not immediately cross-match. 
In this case the coordinates looked similar to SDSSJ1040-1155, and
the stated redshift was identical, $z=5.44$. Additionally the $i$-band magnitude 
was equivalent to SDSS J1040-1155
($i=25.9$ mag). SDSS J1040-1155 is first reported in \citet{Douglas2007} 
and a transcription error seems to have occurred in \citet{Flesch2015}.

\begin{table*}
\begin{tabular}{l r r r r r}
\hline  \hline
Object Name                      & R.A.               & Decl.  &  redshift  & difference \\
                                          & / deg             & / deg  &                & arcsec  \\
\hline
%% QSOs where pos changed (same name)
DELS J1048-0109               & +162.0795417 & -1.1611694        & 6.678   &     12.1 \\
SHELLQs J1429-0002          & +217.3342500 & -0.0353889        & 6.04    &  150.0 \\
DELS J1559+2212               & +239.7878750 & +22.2040083      & 6.07     &     3.0 \\
ATLAS J332.8017-32.1036 & +332.8017500 & -32.1035556      & 6.3394  &    3.2 \\
VIK J2318-3113                  & +349.5764625 & -31.2295417      & 6.4435  &  239.9 \\
\hline  \hline
\end{tabular}
\caption{Five quasars with updated coordinates.}
\label{tab:disagree_objs}
\end{table*}
